---
title: "Guide"
author: "Awesome Me"
date: "2023-05-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load packages
```{r warning = F}
if(!require("pacman")) install.packages("pacman")
pacman::p_load(
  tidyverse,
  janitor,
  readxl,
  knitr,
  ggplot2,
  lubridate,
  arules,
  arulesViz,
  plyr,
  here,
  conflicted,
  RColorBrewer,
  skimr,
  psych,
  reshape2,
  DataExplorer,
  scales,
  onlineretail
)

```



## load dataset

```{r}
# https://archive.ics.uci.edu/ml/machine-learning-databases/00352/Online%20Retail.xlsx
# df <- read_csv(here("data", ".csv"))
# df %>% glimpse()

retail <- as_tibble(onlineretail)
retail %>% glimpse()


```

## inspect dataset structure 

```{r}

retail %>% colnames()  # list of column names
retail %>% nrow()  # how many rows are in data frame?
retail %>% dim()  # dimensions of the data frame?
retail %>% head()  # see the first 6 rows of data frame.  also tail()
retail %>% str()  # see list of columns and data types (numeric, character, etc)
retail %>% summary()  #Statistical summary of data.


```

## examining duplicate records -> for InvoiceNo, StockCode, Description combination

```{r}

duplicated_rows <- retail %>% 
  get_dupes(InvoiceNo, StockCode, Description, InvoiceDate, CustomerID)

duplicated_rows

```

## filter dataset from duplicated rows
```{r}
retail <- retail[!duplicated_rows, ]

```



## remove temporary variables 
```{r}

rm(list = grep("^onlineretail", ls(), value = TRUE))

```

## checking distinct for the whole df or single column
```{r}

n_distinct(retail) 

retail %>% colnames()
  
```

## checking for NAs
```{r}
sum(is.na(retail))

```

## drop all na values (missing values)
```{r}
melbourne_data <- melbourne_data %>% drop_na()

```


## convert column data type
```{r}


```

## delete column(s)
```{r}


```



## normalization
```{r}


```

## visualizing / inspecting descriptive statistics
```{r}
retail %>% gather() %>%
  ggplot(aes(x=value)) + 
  geom_histogram(fill="steelblue", alpha=.7) +
  theme_minimal() +
  facet_wrap(~key, scales="free")

```


## analyze if there are any correlations between variables
```{r}
# density plot
retail %>% plot_density(ncol = 4, ggtheme = theme_minimal())

# correlation plot
corPlot(retail[, 2:7], upper = FALSE, scale = FALSE, main = "Correlation Sales")

```










