---
title: "A Gentle Introduction on Market Basket Analysis — Association Rules"
author: "Awesome Me"
date: "2023-05-10"
link-citations: true
output: html_document
---

### A Gentle Introduction on Market Basket Analysis — Association Rules
#### https://towardsdatascience.com/a-gentle-introduction-on-market-basket-analysis-association-rules-fa4b986a40ce

### dataset 
#### from UCI Machine Learning repository. The dataset is called “Online Retail” and can be found here. It contains all the transactions occurring between 01/12/2010 and 09/12/2011 for a UK-based and registered online retailer

#### https://archive.ics.uci.edu/ml/machine-learning-databases/00352/Online%20Retail.xlsx

### or

#### https://github.com/allanvc/onlineretail/
#### https://cran.r-project.org/web/packages/onlineretail/onlineretail.pdf



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## install packages

```{r}
install.packages("readxl")

install.packages("onlineretail")

```


## load packages
```{r}
pacman::p_load(
  tidyverse,
  readxl,
  knitr,
  ggplot2,
  lubridate,
  arules,
  arulesViz,
  plyr,
  here,
  conflicted,
  onlineretail
)

```


```{r}
# library(pacman)
# 
# library(tidyverse)
# library(readxl)
# library(knitr)
# library(ggplot2)
# library(lubridate)
# library(arules)
# library(arulesViz)
# library(plyr)
# library(here)
# library(conflicted)
# 
# # load data set from onlineretail package
# library(onlineretail)


```

## data preprocessing and exploring

```{r}
# read excel file
# retail <- read_excel(here('data', 'online_retail.xlsx'))
# retail

# rename data set & convert to tibble
retail <- as_tibble(onlineretail)


# inspect dataset structure
retail %>% colnames()  # list of column names
retail %>% nrow()  # how many rows are in data frame?
retail %>% dim()  # dimensions of the data frame?
retail %>% head()  # see the first 6 rows of data frame.  also tail()
retail %>% str()  # see list of columns and data types (numeric, character, etc)
retail %>% summary()  #Statistical summary of data. Mainly for numerics

# remove onlineretail dataset 
rm(list = grep("^onlineretail", ls(), value = TRUE))

```

## clean up and add data to prepare for analysis

```{r}

retail <- retail[complete.cases(retail), ]

retail <- retail %>% mutate(Description = as.factor(Description))
retail <- retail %>% mutate(Country = as.factor(Country))

retail$Date <- as.Date(retail$InvoiceDate)
retail$Time <- format(retail$InvoiceDate,"%H:%M:%S")
retail$InvoiceNo <- as.numeric(as.character(retail$InvoiceNo))

# find columns with NA’s
colSums(is.na(retail)) > 0

```

## count no of customers

```{r}
n_distinct(retail$CustomerID)

```


## What time do people often purchase online?

```{r}
# to find the answer to this question, we need to extract “hour” from the time column.

retail %>% str()

retail$Time <- as.factor(retail$Time)

a <- hms(as.character(retail$Time))

retail$Hour = hour(a)

# remove variable a
rm(list = grep("^a", ls(), value = TRUE))

# count hours
n_distinct(retail$Hour)
retail %>% distinct(Hour)

# visualize rush hours
retail %>% 
  ggplot(aes( x = Hour)) + 
  geom_histogram(stat="count", bins = 15, fill="indianred") + #
  labs(x = "Purchasing Time (for every hour)", y = "No. of People")   


# conclusion 
# There is a clear bias between the hour of day and order volume. Most orders happened between 10:00–15:00
```

## search for all rows with invoice no. 576339 (highest count of row no)

```{r}

retail %>% filter(InvoiceNo == '576339') %>% arrange(desc(Quantity))

```


## count occurrences for each invoice no

```{r}

retail %>% 
  group_by(InvoiceNo) %>% 
  filter(n() > 1) %>% 
  na.omit() %>% # drop rows containing missing values
  summarize(count = n()) %>% 
  arrange(desc(count))

```


## How many items each customer buy?

```{r}
# group by invoice no
items_by_each_customer <- 
  retail %>% 
  group_by(InvoiceNo) %>% 
  summarize(total_item_count = mean(Quantity)) %>% #  sum(Quantity)
  arrange(desc(total_item_count))

items_by_each_customer

ggplot(data = items_by_each_customer, aes(x = total_item_count))+
  geom_histogram(fill="indianred", bins = 100000) + 
  geom_rug()+
  coord_cartesian(xlim=c(0,80))
  
  
# conclusion 
# People mostly purchased less than 10 items (less than 10 items in each invoice)
```

## Top 10 best sellers
```{r}

top_10_best_sellers <- as_tibble(
  retail %>% 
  group_by(StockCode, Description) %>% 
  summarize(count = n()) %>% 
  arrange(desc(count)) #%>% 
  # slice_max(order_by = count, n = 10)
  # top_n(10, count)
)


top_10_best_sellers <- top_10_best_sellers %>% head(10)

top_10_best_sellers %>% str()


# top_10_best_sellers %>%
ggplot(data = top_10_best_sellers, aes(x = reorder(Description, count), y = count))+
  geom_bar(stat="identity",fill="indian red")+
  coord_flip()

```

## transform data frame into transactions
### split data frame it into pieces
```{r}

retail_sorted <- retail[order(retail$CustomerID),]

# split data frame it into pieces based on one or more factors, computes on the pieces, and then returns the results as a data frame
itemList <- ddply(retail,c("CustomerID","Date"), 
                       function(df1)paste(df1$Description, collapse = ","))

```


### keep only transactions

```{r}

# item transactions only needed, so remove CustomerID and Date columns
itemList$CustomerID <- NULL
itemList$Date <- NULL
colnames(itemList) <- c("items")

```

### save the data frame to a csv file and check whether our transaction format is correct.

```{r}

write_csv(itemList, here::here("data", "market-basket.csv"))

```



