---
title: "A Gentle Introduction on Market Basket Analysis — Association Rules"
author: "Awesome Me"
date: "2023-05-10"
link-citations: true
output: html_document
---

### A Gentle Introduction on Market Basket Analysis — Association Rules
#### https://towardsdatascience.com/a-gentle-introduction-on-market-basket-analysis-association-rules-fa4b986a40ce

### dataset 
#### from UCI Machine Learning repository. The dataset is called “Online Retail” and can be found here. It contains all the transactions occurring between 01/12/2010 and 09/12/2011 for a UK-based and registered online retailer

#### https://archive.ics.uci.edu/ml/machine-learning-databases/00352/Online%20Retail.xlsx


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## install packages

```{r}
install.packages("readxl")

```


## load packages

```{r}

library(tidyverse)
library(readxl)
library(knitr)
library(ggplot2)
library(lubridate)
library(arules)
library(arulesViz)
library(here)
# library(plyr)

```

## data preprocessing and exploring

```{r}
retail <- read_excel(here('data', 'online_retail.xlsx'))
retail

retail <- retail[complete.cases(retail), ]

retail <- retail %>% mutate(Description = as.factor(Description))
retail <- retail %>% mutate(Country = as.factor(Country))

retail$Date <- as.Date(retail$InvoiceDate)
retail$Time <- format(retail$InvoiceDate,"%H:%M:%S")
retail$InvoiceNo <- as.numeric(as.character(retail$InvoiceNo))

glimpse(retail)

colnames(retail)
```


## What time do people often purchase online?

```{r}
# In order to find the answer to this question, we need to extract “hour” from the time column.

retail$Time <- as.factor(retail$Time)

a <- hms(as.character(retail$Time))

retail$Time = hour(a)

retail %>% 
  ggplot(aes(x=Time)) + 
  geom_histogram(stat="count",fill="indianred")

# conclusion 
# There is a clear bias between the hour of day and order volume. Most orders happened between 10:00–15:00
```

