---
title: "IBM HR Analytics Employee Attrition & Performance"
author: "Awesome Me"
date: "2023-05-14"
link-citations: true
output: html_document
---

### IBM HR Analytics Employee Attrition & Performance
#### https://www.kaggle.com/datasets/pavansubhasht/ibm-hr-analytics-attrition-dataset/code



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## install h2o
```{r}
# remove any previously installed H2O packages for R.
if ("package:h2o" %in% search()) { detach("package:h2o", unload=TRUE) }

if ("h2o" %in% rownames(installed.packages())) { remove.packages("h2o") }

# download packages that H2O depends on.
pkgs <- c("RCurl","jsonlite")
for (pkg in pkgs) {
  if (! (pkg %in% rownames(installed.packages()))) { install.packages(pkg) }
}

# download and install the H2O package for R
install.packages("h2o", type="source", repos=(c("http://h2o-release.s3.amazonaws.com/h2o/latest_stable_R")))

```


## load packages
```{r}
if(!require("pacman")) install.packages("pacman")
pacman::p_load(
 h2o # initialize H2O 
)

# library(h2o) # initialize H2O 
h2o.init()

```

### set seed for reproducibility
```{r}
# implement or use packages such as bruceR or liver
min_max <- function(x) 
{
  return ((x - min(x)) / (max(x) - min(x)))
}

```



```{r}
# library(pacman)
# 
# library(tidyverse)
# library(readxl)
# library(knitr)
# library(ggplot2)
# library(lubridate)
# library(arules)
# library(arulesViz)
# library(plyr)
# library(here)
# library(conflicted)
# 
# # load data set from onlineretail package
# library(onlineretail)


```

## data preprocessing and exploring

```{r}
# read excel file
# retail <- read_excel(here('data', 'online_retail.xlsx'))
# retail

# rename data set & convert to tibble
retail <- as_tibble(onlineretail)


# inspect dataset structure
retail %>% colnames()  # list of column names
retail %>% nrow()  # how many rows are in data frame?
retail %>% dim()  # dimensions of the data frame?
retail %>% head()  # see the first 6 rows of data frame.  also tail()
retail %>% str()  # see list of columns and data types (numeric, character, etc)
retail %>% summary()  #Statistical summary of data. Mainly for numerics

# remove onlineretail dataset 
rm(list = grep("^onlineretail", ls(), value = TRUE))

```

## clean up and add data to prepare for analysis

```{r}

retail <- retail[complete.cases(retail), ]

retail <- retail %>% dplyr::mutate(Description = as.factor(Description))
retail <- retail %>% dplyr::mutate(Country = as.factor(Country))

retail$Date <- as.Date(retail$InvoiceDate)
retail$Time <- format(retail$InvoiceDate,"%H:%M:%S")
retail$InvoiceNo <- as.numeric(as.character(retail$InvoiceNo))

# find columns with NA’s
colSums(is.na(retail)) > 0

```

## count no of customers

```{r}
n_distinct(retail$CustomerID)

```


## What time do people often purchase online?

```{r}
# to find the answer to this question, we need to extract “hour” from the time column.

retail %>% str()

retail$Time <- as.factor(retail$Time)

a <- hms(as.character(retail$Time))

retail$Hour = hour(a)

# remove variable a
rm(list = grep("^a", ls(), value = TRUE))

# count hours
n_distinct(retail$Hour)
retail %>% distinct(Hour)

# visualize rush hours
retail %>% 
  ggplot(aes( x = Hour)) + 
  geom_histogram(stat="count", bins = 15, fill="indianred") + #
  labs(x = "Purchasing Time (for every hour)", y = "No. of People")   


# conclusion 
# There is a clear bias between the hour of day and order volume. Most orders happened between 10:00–15:00
```

## search for all rows with invoice no. 576339 (highest count of row no)

```{r}

retail %>% filter(InvoiceNo == '576339') %>% arrange(desc(Quantity))

```


## count occurrences for each invoice no

```{r}

retail %>% 
  dplyr::group_by(InvoiceNo) %>% 
  dplyr::filter(n() > 1) %>% 
  na.omit() %>% # drop rows containing missing values
  dplyr::summarize(count = n()) %>% 
  dplyr::arrange(desc(count))

```


## How many items each customer buy?

```{r}
# group by invoice no
items_by_each_customer <- 
  retail %>% 
  dplyr::group_by(InvoiceNo) %>% 
  dplyr::summarize(total_item_count = mean(Quantity)) %>% #  sum(Quantity)
  dplyr::arrange(desc(total_item_count))

items_by_each_customer

ggplot(data = items_by_each_customer, aes(x = total_item_count))+
  geom_histogram(fill="indianred", bins = 100000) + 
  geom_rug()+
  coord_cartesian(xlim=c(0,80))
  
  
# conclusion 
# People mostly purchased less than 10 items (less than 10 items in each invoice)
```

## Top 10 best sellers
```{r}

top_10_best_sellers <- as_tibble(
  retail %>% 
  dplyr::group_by(StockCode, Description) %>% 
  dplyr::summarize(count = n()) %>% 
  dplyr::arrange(desc(count)) #%>% 
  # slice_max(order_by = count, n = 10)
  # top_n(10, count)
)


top_10_best_sellers <- top_10_best_sellers %>% head(10)

top_10_best_sellers %>% str()


# top_10_best_sellers %>%
ggplot(data = top_10_best_sellers, aes(x = reorder(Description, count), y = count))+
  geom_bar(stat="identity",fill="indian red")+
  coord_flip()

```

## transform data frame into transactions
### split data frame it into pieces
```{r}

# retail_sorted <- retail[order(retail$CustomerID),]

# split data frame it into pieces based on one or more factors, computes on the pieces, and then returns the results as a data frame
itemList <- ddply(retail,c("CustomerID","Date"), 
                       function(df1)paste(df1$Description, collapse = ","))

```


### keep only transactions

```{r}

# item transactions only needed, so remove CustomerID and Date columns
itemList$CustomerID <- NULL
itemList$Date <- NULL
colnames(itemList) <- c("items")

```

### save the data frame to a csv file and check whether our transaction format is correct.

```{r}

write_csv(itemList, here::here("data", "market-basket.csv"), quote = "none")

```

### convert data frame to arules transaction data
```{r}
library(here)
retail_transaction <- 
  read.transactions(here::here("data", "market-basket.csv"), format = 'basket', sep = ',')

```

### inspect transaction data
```{r}

retail_transaction %>% str()

retail_transaction %>% summary()

```

## item frequency plot which shows that distribution of the data is right skewed
```{r}
# itemFrequencyPlot(retail_transaction, topN = 20, type = 'absolute')


arules::itemFrequencyPlot(retail_transaction,
   topN=20,
   col=brewer.pal(8,'Pastel2'),
   main='Relative Item Frequency Plot',
   type="relative",
   ylab="Item Frequency (Relative)")

```

## creating some rules 

### all rules have support of at least 0.1% and confidence of at least 80%
```{r}

rules <- apriori(retail_transaction, parameter = list(support = 0.001, confidence = 0.8))

```

### remove redundant rule
```{r}

rules <- rules[!is.redundant(rules)]

```

### sort the rules by decreasing confidence
```{r}
rules <- sort(rules, by = 'confidence', decreasing = TRUE)

rules %>% summary()

```


### inspect the top 10 
```{r}

top_10_rules <- rules[1:10]

top_10_rules %>% inspect()

```




## visualize scatter plot
```{r}

plot(top_10_rules, method = "scatterplot")

```

## visualize interactive scatter plot
```{r}
plot(top_10_rules, measure=c("support", "lift"), shading = "confidence", engine = 'interactive')

# plot(top_10_rules, method = "scatterplot")

```

## visualize graph
```{r}

plot(top_10_rules, method = 'graph')

```

## visualize grouped matrix
```{r}

plot(top_10_rules, method = 'grouped')

```




## interactive scatter plot

```{r}
plot(rules, measure=c("support", "lift"), 
          shading = "confidence",
          engine='interactive')


```

## visualize the relationship between the metrics

### Two-key plot is a scatterplot with shading = "order"
```{r}

plot(top_10_rules, shading="order", control=list(main = "Two-key plot", 
  col=rainbow(5)))

```

### 2D matrix with shading
```{r}
plot(top_10_rules, method="matrix", measure="lift")

plot(top_10_rules, method="matrix", measure="lift", control=list(reorder = "support/confidence"))

```

### 3D matrix
```{r}
plot(top_10_rules, method="matrix3D", measure="lift")

plot(top_10_rules, method="matrix3D", measure="lift", control=list(reorder = "support/confidence"))


```


### matrix with two measures
```{r}

plot(top_10_rules, method="matrix", measure=c("lift", "confidence"))

plot(top_10_rules, method="matrix", measure=c("lift", "confidence"), 
	control=list(reorder = "support/confidence"))

```

### graphs only work well with very few rules
```{r}

## igraph layout generators can be used (see ? igraph::layout_)

# plot(top_10_rules, method="graph", control=list(layout=igraph::layout_in_circle(), igraph::is_igraph()))
# 
# plot(top_10_rules, method="graph", control=list(
#   layout=igraph::with_graphopt(spring.const=5, mass=50)))

```
## parallel coordinates plot
```{r}

plot(top_10_rules, method="paracoord")
plot(top_10_rules, method="paracoord", control=list(reorder=TRUE))

```

<!-- ## Doubledecker plot only works for a single rule -->
<!-- # ```{r} -->
<!-- #  -->
<!-- # oneRule <- sample(rules, 1) -->
<!-- # plot(oneRule, method="doubledecker", data = top_10_rules) -->
<!-- #  -->
<!-- # ``` -->





