---
title: "Migraine and acupuncture"
author: "Awesome Me"
date: "2023-05-07"
output: html_document
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)
```

## (a) What percent of patients in the treatment group were pain free 24 hours after receiving acupuncture?
```{r}
# install openintro package
# install.packages("openintro") # already installed

```
```{r}

# load library
library(tidyverse)
library(openintro)

```

```{r}

# load migraine data set
data(migraine)

migraine %>% head()

migraine %>% distinct(group)

```

```{r}
# percent of patients group (treatment - control)
patients_group_percent <- migraine %>% 
  group_by(group) %>% 
  summarise(count = n()) %>%
  mutate(percent = (round(count / sum(count), 2) * 100)) %>% 
  arrange((percent))

patients_group_percent

```

```{r}
# percent of patients group (treatment - control) & pain free
patients_group_by_pain <- migraine %>% 
  group_by(group, pain_free) %>% 
  summarise(count = n()) %>%
  arrange((group))

patients_group_by_pain

```

```{r}

# Contingency Tables using dplyr & tidyr
# http://analyticswithr.com/contingencytables.html

patients_group_pain_cross_tab <- migraine %>% 
  group_by(group, pain_free) %>% 
  summarise(count = n()) %>%
  spread(pain_free, count) 

patients_group_pain_cross_tab

```

```{r}

# Contingency Tables using table
# http://analyticswithr.com/contingencytables.html

patients_group_pain_table <- table(migraine$group, migraine$pain_free)

patients_group_pain_table

# table(patients_group_pain_table)

```




## (b) What percent were pain free in the control group?

## (c) In which group did a higher percent of patients become pain free 24 hours after receiving acupuncture?

## (d) Your findings so far might suggest that acupuncture is an effective treatment for migraines for all people who suffer from migraines. However this is not the only possible conclusion that can be drawn based on your findings so far. What is one other possible explanation for the observed difference between the percentages of patients that are pain free 24 hours after receiving acupuncture in the two groups?

